animbook: Visualizing changes in performance measures and demographic affiliations using animation

### 2023-10-30 animbook-journal.Rmd
add normalization part for 1st and 2nd scaling

---
title: "animbook: Visualizing changes in performance measures and demographic affiliations using animation"
abstract: >
  Performance is often reported in quantiles or ranks, and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys,", we provide new visualization functions to generate animations that display this type of data, in an R package called \CRANpkg{animbook}. Functions to pre-process the data and prepare it for the animation,  are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
draft: true
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: animbook-journal.bib
---

@@

  For the first and second scaling methods, group splitting is executed using the `quantiles()` and `cut()` functions. The `quantile()` function from the \CRANpkg{stats} package (@stats) takes a numeric vector and outputs the corresponding quantiles to the given probabilities. The output from the `quantile()` function is then used as the `breaks` argument for the `cut()` function that is part of the base R packages (@r).

  In contrast, the third and fourth scaling methods calculate the quantile based on the absolute values scales. To avoid varying values scales among groups, the variables are first normalized to a range between 0 and 1. The default approach then breaks the group equally using the `seq()` function. The `seq()` function takes input values from 0 to 1 and increments by equal steps depending on the number of groups of interest. Additionally, the users have the option to specify the breaks themselves if they choose to do so.

>>>>

---
title: "animbook: Visualizing changes in performance measures and demographic affiliations using animation"
abstract: >
  Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys," we provide new visualization functions to generate animations that display this type of data in an R package called \pkg{animbook}. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
draft: true
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: animbook-journal.bib
---

@@

  For the first and second scaling methods, it is necessary to rank the values based on time, and in cases where a group is provided, they are ranked based on both time and groups. To ensure that the rank scales among different groups are the same, the variables are first normalized to a range between 0 and 1. The third and fourth scaling methods also involve a normalization step, but they are based on raw values instead of rank values.

  All of these scaling methods utilized the `cut()` function from the base R package (@r) to split the values into quantiles. The `cut()` function requires the specification of the `breaks` argument. If it is not provided, the prep function in this package defaults to using the `seq()` function, which sets the minimum and maximum values to 0 and 1, respectively, and then increments by equal steps based on the number of groups of interest. Additionally, the users have the option to specify the breaks themselves if they choose to do so, noting that the breaks provided need to be between 0 and 1, and the length of the vector needs to be the number of groups plus 1.

### 2023-10-30 animbook-journal.tex
add normalization part for 1st and 2nd scaling

\abstract{%
Categorical data has often been presented statically despite its potential to unveil an intriguing movement pattern relevant to the various domains of data, including company ranking quantile in the accounting data and shifts in voter party affiliations. Inspired by the animation in the New York Times article ``Extensive data show punishing reach of racism for black boys,'' this paper aims to develop an R package, animbook, that can generalize the animation to suitably apply to a wider range of data. The package included a function that allowed for the transformation of the data and connection with elements of the plot that would be animated. This results in an animation tool that helps communicate complex data, enhancing the narrative and keeping it engaged for the general audience.
}

@@

For the first and second scaling methods, group splitting is executed using the \texttt{quantiles()} and \texttt{cut()} functions. The \texttt{quantile()} function from the \CRANpkg{stats} package (R Core Team (2013)) takes a numeric vector and outputs the corresponding quantiles to the given probabilities. The output from the \texttt{quantile()} function is then used as the \texttt{breaks} argument for the \texttt{cut()} function that is part of the base R packages (R Core Team (2021)).

In contrast, the third and fourth scaling methods calculate the quantile based on the absolute values scales. To avoid varying values scales among groups, the variables are first normalized to a range between 0 and 1. The default approach then breaks the group equally using the \texttt{seq()} function. The \texttt{seq()} function takes input values from 0 to 1 and increments by equal steps depending on the number of groups of interest. Additionally, the users have the option to specify the breaks themselves if they choose to do so.

@@

\begin{verbatim}
#> # A tibble: 10 x 5
#>    id     time qtile frame color
#>    <fct> <int> <dbl> <dbl> <fct>
#>  1 1      2020     5    11 X    
#>  2 1      2023     5    12 X    
#>  3 2      2020     5    39 X    
#>  4 2      2023     5    40 X    
#>  5 3      2020     5     3 X    
#>  6 3      2023     5     4 X    
#>  7 4      2020     5    12 X    
#>  8 4      2023     5    13 X    
#>  9 5      2020     5    46 X    
#> 10 5      2023     5    47 X
\end{verbatim}

@@

\begin{verbatim}
#> List of 2
#>  $ data    : tibble [400 x 5] (S3: tbl_df/tbl/data.frame)
#>   ..$ id   : Factor w/ 200 levels "1","2","3","4",..: 1 1 2 2 3 3 4 4 5 5 ...
#>   ..$ time : int [1:400] 2020 2023 2020 2023 2020 2023 2020 2023 2020 2023 ...
#>   ..$ qtile: num [1:400] 5 5 5 5 5 5 5 5 5 5 ...
#>   ..$ frame: num [1:400] 11 12 39 40 3 4 12 13 46 47 ...
#>   ..$ color: Factor w/ 2 levels "X","Y": 1 1 1 1 1 1 1 1 1 1 ...
#>  $ settings:List of 7
#>   ..$ gap           : num 0.1
#>   ..$ xbreaks       : int [1:2] 2020 2023
#>   ..$ label         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ order         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ time_dependent: logi FALSE
#>   ..$ runif_min     : num 1
#>   ..$ runif_max     : num 50
#>  - attr(*, "class")= chr "animbook"
\end{verbatim}

@@

From the kangaroo plot (Figure \ref{fig:kan-osiris-figure}), most companies are staying in the same quantile group, with only a small number moving upward or downward. Most of the movement is made by American companies, supporting the OECD report (McGowan, Andrews, and Millot (2017)) that the United States has a higher turnover rate compared to Japan. To emphasize this point, the wallaby plot can be utilized (Figure \ref{fig:osiris-figure}), this reinforces that there are more American companies de-listed compared to Japanese companies. There could be various reasons for this trend, such as the absence of Japanese companies in the Top 25\% in 2006 or the possibility that Japanese companies who started in the Top 25\% in 2006 were not de-listed.

@@

\leavevmode\vadjust pre{\hypertarget{ref-stats}{}}%
R Core Team. 2013. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{http://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-r}{}}%
---------. 2021. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

>>>>

\abstract{%
Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article ``Extensive data show punishing reach of racism for black boys,'' we provide new visualization functions to generate animations that display this type of data in an R package called \pkg{animbook}. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
}

@@

For the first and second scaling methods, it is necessary to rank the values based on time, and in cases where a group is provided, they are ranked based on both time and groups. To ensure that the rank scales among different groups are the same, the variables are first normalized to a range between 0 and 1. The third and fourth scaling methods also involve a normalization step, but they are based on raw values instead of rank values.

All of these scaling methods utilized the \texttt{cut()} function from the base R package (R Core Team (2021)) to split the values into quantiles. The \texttt{cut()} function requires the specification of the \texttt{breaks} argument. If it is not provided, the prep function in this package defaults to using the \texttt{seq()} function, which sets the minimum and maximum values to 0 and 1, respectively, and then increments by equal steps based on the number of groups of interest. Additionally, the users have the option to specify the breaks themselves if they choose to do so, noting that the breaks provided need to be between 0 and 1, and the length of the vector needs to be the number of groups plus 1.

@@

\begin{verbatim}
#> # A tibble: 10 x 5
#>    id     time qtile frame color
#>    <fct> <int> <dbl> <dbl> <fct>
#>  1 1      2020     5     2 X    
#>  2 1      2023     5     3 X    
#>  3 2      2020     5    20 X    
#>  4 2      2023     5    21 X    
#>  5 3      2020     5    34 X    
#>  6 3      2023     5    35 X    
#>  7 4      2020     5    20 X    
#>  8 4      2023     5    21 X    
#>  9 5      2020     5    45 X    
#> 10 5      2023     5    46 X
\end{verbatim}

@@

\begin{verbatim}
#> List of 2
#>  $ data    : tibble [400 x 5] (S3: tbl_df/tbl/data.frame)
#>   ..$ id   : Factor w/ 200 levels "1","2","3","4",..: 1 1 2 2 3 3 4 4 5 5 ...
#>   ..$ time : int [1:400] 2020 2023 2020 2023 2020 2023 2020 2023 2020 2023 ...
#>   ..$ qtile: num [1:400] 5 5 5 5 5 5 5 5 5 5 ...
#>   ..$ frame: num [1:400] 2 3 20 21 34 35 20 21 45 46 ...
#>   ..$ color: Factor w/ 2 levels "X","Y": 1 1 1 1 1 1 1 1 1 1 ...
#>  $ settings:List of 7
#>   ..$ gap           : num 0.1
#>   ..$ xbreaks       : int [1:2] 2020 2023
#>   ..$ label         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ order         : chr [1:5] "A" "B" "C" "D" ...
#>   ..$ time_dependent: logi FALSE
#>   ..$ runif_min     : num 1
#>   ..$ runif_max     : num 50
#>  - attr(*, "class")= chr "animbook"
\end{verbatim}

@@

From the kangaroo plot (Figure \ref{fig:kan-osiris-figure}), most companies are staying in the same quantile group, with only a small number moving upward or downward. Most of the movement is made by American companies, supporting the OECD report (McGowan, Andrews, and Millot (2017)) that the United States has a higher turnover rate compared to Japan. To emphasize this point, the wallaby plot can be utilized (Figure \ref{fig:osiris-figure}). There are clearly more American companies de-listed compared to Japanese companies, and we can see that there are no Japanese companies in the Top 25\% in 2006 that were de-listed.

@@

\leavevmode\vadjust pre{\hypertarget{ref-r}{}}%
R Core Team. 2021. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
